---
title: "Development"
author: "Anna Prohofsky"
date: "2024-07-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) #for knitting the PDF
if (!require("ggplot2")) install.packages("ggplot2"); library(ggplot2) #for pretty plots
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse) #a collection of different packages for data science, such as ggplot and dplyr. ALWAYS DO LAST!! 
```

## Read in data & check packaging

```{r reading in data}
hatchling_data <- read.csv("hatchling_data.csv") 

head(hatchling_data)
tail(hatchling_data)
summary(hatchling_data)
str(hatchling_data)

```
Head/tail match the .csv file. The str looks fine. days_incubating and harrison_stage were both read in as integers, and substrate as character, so we're good there.I realize that I misspelled interocular as "interoccular" but I fixed it for the figures.

Let's filter out that one outlier, the one at Harrison stage 32. It hatched dead and underdeveloped, so I think we can consider it a loss like the ones lost to water mold.

```{r Filter out N30}
hatchling_data_filtered <- hatchling_data %>% filter(id != 'N30')
tail(hatchling_data_filtered)
```


## T-tests

Before we can do t-tests, we need to confirm that the data has normal distribution. Let's do some shapiro tests

```{r confirm normal distribution}
shapiro.test(hatchling_data_filtered$days_incubating)
shapiro.test(hatchling_data_filtered$body_length)
shapiro.test(hatchling_data_filtered$interoccular_distance)
shapiro.test(hatchling_data_filtered$gape_width)
shapiro.test(hatchling_data_filtered$growth_rate)

```
Looks like body length and interocular distance are normally distributed and days incubating and growth rate aren't. Meanwhile, gape width is on the cusp. Let's zoom in on the gape width and take a look at a Q-Q plot

```{r gape width Q-Q plot}
ggplot(data = hatchling_data_filtered, aes(sample = gape_width)) +
  geom_qq() +
  geom_qq_line() +
  ggtitle("Q-Q Plot for Gape Width")
```
Hmmm....I'm on the fence still. What does the histogram look like?

```{r gape width histogram}
ggplot(data = hatchling_data_filtered, aes(x = gape_width)) +
  geom_histogram(binwidth = 0.01, fill = "blue", color = "black") +
  ggtitle("Histogram of Gape Width")
```
If you squint you can see a curve, but even then it's shifted to the left with a long tail to the right. Despite the shapiro test p-value = 0.05175 which is technically above 0.05, I'm going to call this not normally distributed. 

This means our normally distributed data is body length with p-value = 0.9855 and interocular distance with p-value = 0.2457. Let's move forward with the t-test there.

```{r t tests}
t.test(body_length ~ substrate, data = hatchling_data_filtered)
t.test(interoccular_distance  ~ substrate, data = hatchling_data_filtered)
```
Significance in both!

## Non-paramentric tests

For the other data, we can use the Wilcoxon rank-sum test.

```{r Wilcoxon rank-sum}
wilcox.test(days_incubating ~ substrate, data = hatchling_data_filtered)
wilcox.test(gape_width ~ substrate, data = hatchling_data_filtered)
wilcox.test(growth_rate ~ substrate, data = hatchling_data_filtered)
```
And three more results with statistical significance, yay!

## Stat Reporting
```{r means and SD}
summary_table <- hatchling_data_filtered %>%
  group_by(substrate) %>%
  summarise(across(where(is.numeric), list(mean = ~mean(.), sd = ~sd(.)), .names = "{.col}_{.fn}"))

summary_table
```


## Figures

Let's construct some figures now. I think box plots are appropriate

```{r incubation figure}
ggplot(data=hatchling_data_filtered, aes(x=substrate, y=days_incubating, fill=substrate)) +
  geom_boxplot(width=0.25) +
  labs(x = "Wrapping Material", y = "Days Incubating") +
  theme_minimal() +
  scale_fill_manual(values = c("P" = "gray70", "L" = "gray30")) + 
  scale_x_discrete(labels = c("P" = "Live Plant", "L" = "Dead Leaf")) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        legend.position = "none")

```

```{r body length figure}
ggplot(data=hatchling_data_filtered, aes(x=substrate, y=body_length, fill=substrate)) +
  geom_boxplot(width=0.25) +
  labs(x = "Wrapping Material", y = "Body Length (mm)") +
  theme_minimal() +
  scale_fill_manual(values = c("P" = "gray70", "L" = "gray30")) + 
  scale_x_discrete(labels = c("P" = "Live Plant", "L" = "Dead Leaf")) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        legend.position = "none")
```

```{r interoc figure}
ggplot(data=hatchling_data_filtered, aes(x=substrate, y=interoccular_distance, fill=substrate)) +
  geom_boxplot(width=0.25) +
  labs(x = "Wrapping Material", y = "Interocular Distance (mm)") +
  theme_minimal() +
  scale_fill_manual(values = c("P" = "gray70", "L" = "gray30")) + 
  scale_x_discrete(labels = c("P" = "Live Plant", "L" = "Dead Leaf")) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        legend.position = "none")
```

```{r}
ggplot(data=hatchling_data_filtered, aes(x=substrate, y=gape_width, fill=substrate)) +
  geom_boxplot(width=0.25) +
  labs(x = "Wrapping Material", y = "Gape Width (mm)") +
  theme_minimal() +
  scale_fill_manual(values = c("P" = "gray70", "L" = "gray30")) + 
  scale_x_discrete(labels = c("P" = "Live Plant", "L" = "Dead Leaf")) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        legend.position = "none")
```

```{r}
ggplot(data=hatchling_data_filtered, aes(x=substrate, y=grow, fill=substrate)) +
  geom_boxplot(width=0.25) +
  labs(x = "Wrapping Material", y = "Gape Width (mm)") +
  theme_minimal() +
  scale_fill_manual(values = c("P" = "gray70", "L" = "gray30")) + 
  scale_x_discrete(labels = c("P" = "Live Plant", "L" = "Dead Leaf")) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        legend.position = "none")
```

## Reference

1: ChatGPT assisted with occasional code error questions