---
title: "Development"
author: "Anna Prohofsky"
date: "2024-07-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) #for knitting the PDF

if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse) #a collection of different packages for data science, such as ggplot and dplyr. ALWAYS DO LAST!! 
```

## Read in data & check packaging

```{r reading in data}
hatchling_data <- read.csv("hatchling_data.csv") 

head(hatchling_data)
tail(hatchling_data)
summary(hatchling_data)
str(hatchling_data)

```
Head/tail match the .csv file. The str looks fine. days_incubating and harrison_stage were both read in as integers, and substrate as character, so we're good there.

## T-tests

Let's start out with t tests for each dependent variable
```{r t tests}
t.test(days_incubating ~ substrate, data = hatchling_data)
t.test(body_length ~ substrate, data = hatchling_data)
t.test(interoccular_distance ~ substrate, data = hatchling_data)
t.test(gape_width ~ substrate, data = hatchling_data)
t.test(harrison_stage ~ substrate, data = hatchling_data)
```
Looks like we have significance in days_incubating (p = 0.001641) but nothing else. Gape width is pretty darn close at 0.06. 

I wonder what happens when we remove the underdeveloped larva? N30 only made it to Harrison stage 32 before dying. 

```{r filter out N30}
hatchling_data_filtered <- hatchling_data %>% filter(id != 'N30')
tail(hatchling_data_filtered)
```

With N30 successfully removed, let's rerun the t-tests, excluding the harrison_stage since all of the remaining ones are at 40.

```{r rerun t tests}
t.test(days_incubating ~ substrate, data = hatchling_data_filtered)
t.test(body_length ~ substrate, data = hatchling_data_filtered)
t.test(interoccular_distance ~ substrate, data = hatchling_data_filtered)
t.test(gape_width ~ substrate, data = hatchling_data_filtered)
```
Interesting, now we have significance in days_incubating (p-value = 0.002431), body length (p-value = 0.00947), and gape width (p-value = 0.002533)

There's likely some interplay here between the dependent variables. It makes sense that the longer the eggs incubate, the larger the larvae may be, and larger larvae would have larger mouths.  

## Reference

ChatGPT assisted with occasional code error questions