---
title: "predation"
author: "Anna Prohofsky"
date: "2024-09-06"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) #for knitting the PDF
if (!require("ggplot2")) install.packages("ggplot2"); library(ggplot2) #for pretty plots
if (!require("rstatix")) install.packages("rstatix"); library(rstatix) #needed for the Dunn test
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse) #a collection of different packages for data science, such as ggplot and dplyr. ALWAYS DO LAST!! 
```

## Read in data & check packaging
```{r reading in data}
predation_data <- read.csv("predation_data.csv") 

head(predation_data)
tail(predation_data)
summary(predation_data)
str(predation_data)
```
Everything looks fine, except it did not preserve my significant digits, though I don't think that matters. 

Next let's check for normality & homogeneity

## Assumptions Check

```{r normalcy check}
predation_data %>%
  group_by(Substrate) %>%
  summarise(shapiro_test = shapiro.test(Eggs.Eaten)$p.value)
```
Looks like N is normally distributed, P isn't, and L is on the cusp, so a non-parametric test is needed. 

## Statistical tests

We'll start with a Kruskal-Wallis test to determine if there is any significance
```{r kw test}
kruskal.test(Eggs.Eaten ~ Substrate, data = predation_data)
```
Looks like there is significant difference there, though between what groups is the question.

Next a post-hoc test of some kind is needed. I'll try a Dunntest with a Bonferroni correction since I have 3 groups.
```{r Dunn test}
dunn_result <- dunn_test(formula = Eggs.Eaten ~ Substrate, 
                         data = predation_data, 
                         p.adjust.method = "bonferroni")

dunn_result
```
So there's statistical significance between Leaf and None, but not between the others.

## Figure

```{r boxplot}
ggplot(data=predation_data, aes(x=Substrate, y=Eggs.Eaten, fill=Substrate)) +
  geom_boxplot(width=0.25) +
  labs(x = "Wrapping Material", y = "Percentage Eggs Eaten") +
  theme_minimal() +
  scale_fill_manual(values = c("P" = "gray70", "L" = "gray30", "N" = "gray10")) + 
  scale_x_discrete(labels = c("P" = "Live Plant", "L" = "Dead Leaf", "N" = "None")) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        legend.position = "none") +
geom_text(
    aes(x = 1, y = 0.85, label = "a"), 
    size = 5, vjust = -0.5) +  
  geom_text(
    aes(x = 2, y = 0.85, label = "b"), 
    size = 5, vjust = -0.5) +  
  geom_text(
    aes(x = 3, y = 0.85, label = "ab"), 
    size = 5, vjust = -0.5)
```


## Reference

1: ChatGPT assisted with occasional code error questions
